---
title: "Day 6 EDS 221 Task"
author: "Dalila Lara"
date: "2022-08-15"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(here)
library(tidyverse)
library(janitor)
```
# Task 2 SBC Lobsters 

Read data in, note na values, update line for na values, convert names to lower case, convert using uncount()
```{r}
#uncount() gives each observation its own row
lobsters <- read_csv(here("data","Lobster_Abundance_All_Years_20210412.csv"), na = c("-99999", "")) %>% 
  clean_names() %>% 
  uncount(count)
```

Counts and mean sizes by site & year
```{r}
#makes table to summarize data 
lobsters_c_yr <- lobsters |>
  group_by(site, year) |>
  summarize(mean_length = mean(size_mm),
            counts = n())
  

#make graph of summarized data 
ggplot(data = lobsters_c_yr, 
       aes(x = year, y = counts)) +
  geom_line(aes(color = site)) #+
  #facet_wrap(~site) +

```

Create a subset of 2020 data
```{r}
lobsters2020 <- lobsters |>
  filter(year == 2020)
```

Find counts of lobster at each site group_by()
legal "yes" "no"
```{r}
lobsters_c_2020 <- lobsters |>
  filter(year == 2020) |>
  mutate(legal = case_when(size_mm >= 79.76 ~ "Yes", TRUE ~ "No")) |>
  group_by(site, legal) |>
  summarize(counts = n())
```

Create stacked column graph
```{r}
ggplot(data = lobsters_c_2020, aes( x = site, y = counts)) +
  geom_col(position = "fill",
    aes(fill = legal)) 
```
Which two sites had the largest proportion of legal lobsters in 2020? Explore the metadata to come up with a hypothesis about why that might be.

  In 2020 site NAPL & IVEE site had the largest proportion of legal lobsters.
  
## Task 3 Random lobster wrangling
filter() practice

```{r}
# Subset with data from three specified sites
ex_a <- lobsters |>
  filter(site == c("IVEE", "CARP", "NAPL"))

unique(ex_a$site)
```


```{r}
# Subset with only lobsters in august
ex_b <- lobsters |>
  filter(month == 8)
```

```{r}
# Subset with lobsters @ AQUE or carspace length > 70mm
ex_c <- lobsters |>
  filter(site == "AQUE" | size_mm > 70)
```

```{r}
# subset with no NAPL as site
ex_d <- lobsters |>
  filter(!site %in% "NAPL")

#check 
unique(ex_d$site)
```

```{r}
#group_by() %>% summarize() practice
ex_e <- lobsters |>
  group_by(size_mm, site) |>
  summarize(mean(size_mm, na.rm = TRUE),
            sd(size_mm, na.rm = TRUE))
```

```{r}
#find max carspace length by site and month
ex_f <- lobsters |>
  group_by(site, month) |>
  summarize(max(size_mm))
```

```{r}
#mutate practice 
# new column that contain length converted to centimeters 
ex_g <- lobsters |>
  mutate(size_cm = size_mm * 100 * 1000) 

#site column to lower case
ex_h <- lobsters |>
  mutate(site = tolower(site))
  
#convert area column to character 
ex_i <- lobsters |> 
  mutate(area = as.character(area))
```


